---
output: 
  html_document:
    df_print: paged
    code_folding: hide
    code_download: true
    number_sections: true
    toc: true
    toc_float: true
params:
  title: "integration analysis report"
  sample_table: "sample_table_ss.txt"
  n_cols: 2
  input_object: "out_rds/sobj.separate.cc.rds"
  object_dir: "out_rds/"
  integration_dims: 20
  integration_npcs: 20
  integration_resolution: 1.0
  evaluate_reproducibility_on: "sample"
title: |
  | CUBI SCBA Single Cell Pipeline
  | 
  | `r params$title`
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, cache=FALSE, cache.lazy=FALSE, message=FALSE, warning=FALSE)

suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(gtools))
suppressPackageStartupMessages(library(ggplot2))

options(future.globals.maxSize = 12 * 1024 * 1024^2)

sample_df <- read.delim(params$sample_table, header = T)
sample_df$library <- as.character(sample_df$library)
sample_df$sample <- as.character(sample_df$sample)


if (any(is.na(unique(sample_df$multiplex)) | unique(sample_df$multiplex)=="false" | unique(sample_df$multiplex) == FALSE )){
	multiplex_condition <- FALSE
} else if (length(unique(sample_df$multiplex)) > 1){
	multiplex_condition <- TRUE
}	else {
	stop("Something is wrong with the multiplex setup. Please modify sample table.", call. = FALSE)
}

```

# Analysis Parameters
For this Seurat analysis, we used the following parameters:

```{r print parameters}

data.frame(parameter=names(params), value=sapply(names(params), function(x) params[[x]])) %>%
  filter(parameter != "title") %>%
  tibble::remove_rownames()

```

First, read in the seurat object.

```{r get_data,fig.width=8,fig.height=4}

possible_input_objects <- c("sobj.rds", "sobj.separate.cc.rds", "sobj.diff.cc.rds",
                            "sobj.df.rds", "sobj.df.separate.cc.rds", "sobj.df.diff.cc.rds")

if (params$input_object %in% paste0(params$object_dir,possible_input_objects)){
  sobj <- readRDS(params$input_object)
    } else {
        stop("Please decide on object (sobj.rds|sobj.separate.cc.rds|sobj.diff.cc.rds|sobj.df.rds|sobj.df.separate.cc.rds|sobj.df.diff.cc.rds)", call. = FALSE)
}
    
DefaultAssay(sobj) <- "RNA"
```

# Data integration comparisons
## Keep previous variable features 

We explore the effect of different methods of splitting the data (e.g. by sample, condition, batch) on the data integration results. What works best for removal of batch effects?

```{r loop_splits, results = FALSE}

splits <- colnames(sample_df)
splits <- splits[splits %in% colnames(sobj[[]])]
if (!multiplex_condition){
  splits <- splits[splits != "multiplex"]
}

check_splits <- function(split){
  length(unique(sample_df[,split]))>=2
}

keep_splits <- sapply(splits, check_splits)
splits <- splits[keep_splits]

if (c("library") %in% splits & length(unique(sample_df$library)) == length(unique(sample_df$sample))){
  keep <- which(setdiff(splits, "library") %in% splits)
  splits <- splits[keep]
}

if (is.null(splits)){
  stop("The names of the splitting conditions do not match in sample table and metadata of the object. Please correct this.", call. = FALSE)
} else {
  
  test_splitting_integration <- function(split){
    sobj.spl <- SplitObject(sobj, split.by=split)
    n_cells_per_object <- lapply(names(sobj.spl), function(x) length(Cells(sobj.spl[[x]])))
    if (any(n_cells_per_object <= 150)){
      warning(paste0("Using integration on: ", split," these objects have less than 150 cells and were removed before continuing: ", names(sobj.spl[n_cells_per_object <= 150])))
      sobj.spl <- sobj.spl[n_cells_per_object > 150]
    }
    anchors <- FindIntegrationAnchors(sobj.spl, dims=1:params$integration_dims, verbose=FALSE, reduction = "rpca")
    sobj.integrated <- IntegrateData(anchors, dims=1:params$integration_dims, verbose=FALSE) 
    DefaultAssay(sobj.integrated) <- "integrated"
    sobj.integrated <- ScaleData(sobj.integrated, verbose = FALSE) %>%
      RunPCA(npcs = params$integration_npcs, verbose = FALSE) %>%
      FindNeighbors(dims=1:params$integration_dims, verbose = FALSE) %>%
      RunUMAP(dims = 1:params$integration_dims, verbose = FALSE) %>%
      FindClusters(resolution=params$integration_resolution, verbose=FALSE)
  }
  integrated <- lapply(splits, test_splitting_integration)
}

names(integrated) <- splits

```


```{r plot integrated, eval = (length(unique(sample_df$sample)) <= 12), fig.width=9, fig.height=5, results = F} 
plot_integrated <- function(integrated_object){
  DimPlot(integrated[[integrated_object]], reduction = "umap", label = F, split.by = "sample", ncol = params$n_cols) + ggtitle(paste0("integration on: ", integrated_object))
}

lapply(names(integrated), plot_integrated)

```

```{r plot integrated many samples, eval = (length(unique(sample_df$sample)) <= 30 & length(unique(sample_df$sample)) > 12), fig.width=20, fig.height=20, results = F} 
plot_integrated <- function(integrated_object){
  DimPlot(integrated[[integrated_object]], reduction = "umap", label = F, split.by = "sample", ncol = params$n_cols) + ggtitle(paste0("integration on: ", integrated_object))
}

lapply(names(integrated), plot_integrated)

```



## Recompute variable features 
Some variable features might have been removed from the input object due to cell/doublet filtering. We therefore recompute the variable features before integration and explore how this affects the results of data integration.

```{r force_meta_features, fig.width=8, fig.height=5, results = FALSE}

sobj.forced <- sobj

sobj.forced@assays$RNA[['scale.data']] <- NULL
sobj.forced@assays[['RNA']]@var.features <- character(0)
sobj.forced@assays$RNA[['meta.features']] <- NULL

if (is.null(splits)){
  stop("The names of the splitting conditions do not match in sample table and metadata of the object. Please correct this.", call. = FALSE)
} else {
  
  test_splitting_integration_forced <- function(split){
    sobj.spl <- SplitObject(sobj.forced, split.by=split)
    n_cells_per_object <- lapply(names(sobj.spl), function(x) length(Cells(sobj.spl[[x]])))
    if (any(n_cells_per_object <= 150)){
      warning(paste0("Using integration on: ", split," these objects have less than 150 cells and were removed before continuing: ", names(sobj.spl[n_cells_per_object <= 150])))
      sobj.spl <- sobj.spl[n_cells_per_object > 150]
    }
    anchors <- FindIntegrationAnchors(sobj.spl, dims=1:params$integration_dims, verbose=FALSE, reduction = "rpca")
    sobj.integrated <- IntegrateData(anchors, dims=1:params$integration_dims, verbose=FALSE) 
    DefaultAssay(sobj.integrated) <- "integrated"
    sobj.integrated <- ScaleData(sobj.integrated, verbose = FALSE) %>%
      RunPCA(npcs = params$integration_npcs, verbose = FALSE) %>%
      FindNeighbors(dims=1:params$integration_dims, verbose = FALSE) %>%
      RunUMAP(dims = 1:params$integration_dims, verbose = FALSE) %>%
      FindClusters(resolution=params$integration_resolution, verbose=FALSE)
  }
  integrated.forced <- lapply(splits, test_splitting_integration_forced)
}

names(integrated.forced) <- splits

```

```{r plot integrated forced, eval = (length(unique(sample_df$sample)) <= 12), fig.width=9, fig.height=5, results = F} 
plot_integrated_forced <- function(integrated_object){
  DimPlot(integrated.forced[[integrated_object]], reduction = "umap", label = F, split.by = "sample", ncol = params$n_cols) + ggtitle(paste0("integration on: ", integrated_object))
}

lapply(names(integrated.forced), plot_integrated_forced)

```

```{r plot integrated many samples forced, eval = (length(unique(sample_df$sample)) <= 30 & length(unique(sample_df$sample)) > 12), fig.width=20, fig.height=20, results = F} 
plot_integrated_forced <- function(integrated_object){
  DimPlot(integrated.forced[[integrated_object]], reduction = "umap", label = F, split.by = "sample", ncol = params$n_cols) + ggtitle(paste0("integration on: ", integrated_object))
}

lapply(names(integrated.forced), plot_integrated_forced)

```


## Data integration results
### Percentage of cells per cluster

```{r pct cells per cluster, fig.width=12,fig.height=10, eval = params$evaluate_reproducibility_on =="sample", results=F}

plot_cluster_percentages <- function(integrated_object){
  integrated[[integrated_object]][[]] %>%
  dplyr::select(sample, seurat_clusters) %>%
  group_by(sample, seurat_clusters) %>%
  summarise(cell_freq = n()) %>%
  ungroup() %>%
  group_by(sample) %>%
  mutate(cells_in_sample = sum(cell_freq), perc_in_cluster = cell_freq/sum(cell_freq)*100) %>%
  ggplot(aes(sample, perc_in_cluster, fill = seurat_clusters)) + geom_bar(position = "stack", stat = "identity") +
  theme(legend.position = "bottom", axis.text.x = element_text(angle=90))+
  ylab("% cells in cluster") + ggtitle(paste0("integration on: ",integrated_object))
}

lapply(names(integrated), plot_cluster_percentages)


plot_cluster_percentages_forced <- function(integrated_object){
  integrated.forced[[integrated_object]][[]] %>%
  dplyr::select(sample, seurat_clusters) %>%
  group_by(sample, seurat_clusters) %>%
  summarise(cell_freq = n()) %>%
  ungroup() %>%
  group_by(sample) %>%
  mutate(cells_in_sample = sum(cell_freq), perc_in_cluster = cell_freq/sum(cell_freq)*100) %>%
  ggplot(aes(sample, perc_in_cluster, fill = seurat_clusters)) + geom_bar(position = "stack", stat = "identity") +
  theme(legend.position = "bottom", axis.text.x = element_text(angle=90))+
  ylab("% cells in cluster") + ggtitle(paste0("forced integration on: ",integrated_object))
}

lapply(names(integrated.forced), plot_cluster_percentages_forced)

```

```{r pct cells per cluster replicates, fig.width=12,fig.height=5, eval = params$evaluate_reproducibility_on !="sample", results = F}

plot_cluster_percentages <- function(integrated_object){
  integrated[[integrated_object]][[]] %>%  
    dplyr::select(sample, all_of(params$evaluate_reproducibility_on), seurat_clusters) %>%
    dplyr::group_by_all() %>%
    dplyr::summarise(cell_freq=n()) %>%
    ungroup() %>%
    group_by(sample) %>%
    mutate(cells_in_sample=sum(cell_freq), perc_in_cluster=cell_freq/sum(cell_freq)*100) %>%
    ggplot(aes(seurat_clusters, perc_in_cluster, fill=.data[[params$evaluate_reproducibility_on]]))+
    geom_boxplot(position=position_dodge(width=.5),width=.5, outlier.shape = 1, outlier.colour = "red")+
    geom_point(position=position_dodge(width=.5),colour='grey30', shape=1, size=2)+
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))+
    xlab("")+ylab("% cells_in_sample") + 
    ggtitle(paste0("integration on: ",integrated_object))
}  

lapply(names(integrated), plot_cluster_percentages)


plot_cluster_percentages_forced <- function(integrated_object){
  integrated.forced[[integrated_object]][[]] %>%  
    dplyr::select(sample, all_of(params$evaluate_reproducibility_on), seurat_clusters) %>%
    dplyr::group_by_all() %>%
    dplyr::summarise(cell_freq=n()) %>%
    ungroup() %>%
    group_by(sample) %>%
    mutate(cells_in_sample=sum(cell_freq), perc_in_cluster=cell_freq/sum(cell_freq)*100) %>%
    ggplot(aes(seurat_clusters, perc_in_cluster, fill=.data[[params$evaluate_reproducibility_on]]))+
    geom_boxplot(position=position_dodge(width=.5),width=.5, outlier.shape = 1, outlier.colour = "red")+
    geom_point(position=position_dodge(width=.5),colour='grey30', shape=1, size=2)+
    theme(axis.text.x = element_text(angle=45,vjust=1,hjust=1))+
    xlab("")+ylab("% cells_in_sample") + 
    ggtitle(paste0("forced integration on: ",integrated_object))
}  
lapply(names(integrated.forced), plot_cluster_percentages_forced)

```

Finally, store Seurat objects.

```{r write_rds, results = FALSE}

dir.create(params$object_dir, recursive = T, showWarnings = F)
lapply(names(integrated), function(x) saveRDS(integrated[[x]], paste0(params$object_dir, "/sobj.",x,".rds" )))
lapply(names(integrated.forced), function(x) saveRDS(integrated.forced[[x]], paste0(params$object_dir, "/sobj.",x,".forced.rds" )))

```

# Appendix

## Plot all samples after integration
If more than 20 samples are present in the analysis, plot them separately here.

```{r plot integrated too many samples, eval = (length(unique(sample_df$sample)) >= 20), fig.width=6, fig.height=3, results = F}
plot_integrated <- function(integrated_object){
   relevant_objects <- integrated[[integrated_object]] %>%
     SplitObject(split.by = "sample") %>%
     lapply(function(x) nrow(x[[]]) >1) %>%
     unlist() %>%
     which()
   
     SplitObject(integrated[[integrated_object]], split.by = "sample")[relevant_objects] %>%
     lapply(function(x) x %>%
              DimPlot(reduction = "umap", label = F, group.by="sample") +
              ggtitle(paste0("integration on: ",integrated_object)))
}

lapply(names(integrated), plot_integrated)

```

Forced results for more than 20 samples.

```{r plot integrated too many samples forced, eval = (length(unique(sample_df$sample)) >= 20), fig.width=6, fig.height=3, results = F}

plot_integrated_forced <- function(integrated_object){
   relevant_objects <- integrated.forced[[integrated_object]] %>%
     SplitObject(split.by = "sample") %>%
     lapply(function(x) nrow(x[[]]) >1) %>%
     unlist() %>%
     which()
   
     SplitObject(integrated.forced[[integrated_object]], split.by = "sample")[relevant_objects] %>%
     lapply(function(x) x %>%
              DimPlot(reduction = "umap", label = F, group.by="sample") +
              ggtitle(paste0("integration on: ",integrated_object)))
}

lapply(names(integrated), plot_integrated_forced)

```

```{r sessionInfo}

sessionInfo()
```
